---
title: "NFL 2023 QB Analysis"
author: "Andre Contreras"
date: "2023-11-20"
output: prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(readxl)
library(ggrepel)
library(readr)
```

# NFL 2023 QB Performance Project


## Overall QB Performance:

```{r, fig.width=12, warning=FALSE, message=FALSE}
qbdata <- read_csv("C:\\Users\\ajcon\\Downloads\\Portfolio\\qbdata.csv")

filtered_qbdata <- qbdata %>% 
  filter(TotSnaps > 100)

mvp_finalists <- filtered_qbdata %>% 
  top_n(5, filtered_qbdata$MVP_Rating)

ggplot(filtered_qbdata, aes(x = Efficiency, y = Grade, col = Player)) +
  geom_point() +
  geom_label_repel(data = mvp_finalists, aes(label = Player), nudge_x = -0.001, nudge_y = -0.05, segment.color = "black", segment.size = 0.5) +
  labs(x = "Weighted Efficiency Scale", y = "Grading Scale", title = "Overall QB Performance 2023-2024 NFL Season") +
  theme_minimal() +
  theme(legend.position = "right")

```

> As we can see, my MVP rating suggests that the 5 MVP Finalists be the named above. When comparing it to the real life finalists, only Jared Goff was not included (replaced by Christian McCaffrey).

> Note: In case you were wondering why the 2nd and 3rd highest Grade datapoints (Jake Browning & Kirk Cousins) were not considered an MVP candidate, it is because the formula that quantifies a player's mvp rating incorporates the player's total amount of snaps, and wins. Both of these players played less than half a season. 

## QB Performance by Team

```{r, fig.width=20, fig.height=15}
ggplot(filtered_qbdata, aes(x = Efficiency, y = Grade)) +
  geom_point() +
  geom_text(aes(label = Player), nudge_x = 0, nudge_y = -0.05) +
  facet_wrap(. ~ Team) +
  labs(x = "Efficiency", y = "Grade", title = "Quarterback Performance by Team")
```

