---
title: "NFL 2023 QB Analysis"
author: "Andre Contreras"
date: "2023-11-20"
output: prettydoc::html_pretty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(readxl)
library(ggrepel)
library(readr)
library(caret)
library(fmsb) # Radar Graph
```

# NFL 2023 QB Performance Project


## Overall QB Performance:

```{r, fig.width=12, warning=FALSE, message=FALSE}
qbdata <- read_excel("C:\\Users\\ajcon\\Downloads\\Portfolio\\NFL Projects\\qbdata.xlsx")

filtered_qbdata <- qbdata %>% 
  filter(TotSnaps > 100)

mvp_finalists <- filtered_qbdata %>% 
  top_n(5, filtered_qbdata$MVP_Rating)

ggplot(filtered_qbdata, aes(x = Efficiency, y = Grade, col = Player)) +
  geom_point() +
  geom_label_repel(data = mvp_finalists, aes(label = Player), nudge_x = -0.001, nudge_y = -0.05, segment.color = "black", segment.size = 0.5) +
  labs(x = "Weighted Efficiency Scale", y = "Grading Scale", title = "Overall QB Performance 2023-2024 NFL Season") +
  theme_minimal() +
  theme(legend.position = "right")

# make a linear regression model to predict the number of mvp points a qb gets, whoever has the most points wins mvp
```

> As we can see, my MVP rating suggests that the 5 MVP Finalists be the named above. When comparing it to the real life finalists, only Jared Goff was not included (replaced by Christian McCaffrey).

> Note: In case you were wondering why the 2nd and 3rd highest Grade datapoints (Jake Browning & Kirk Cousins) were not considered an MVP candidate, it is because the formula that quantifies a player's mvp rating incorporates the player's total amount of snaps, and wins. Both of these players played less than half a season. 

## QB Performance by Team

```{r, fig.width=20, fig.height=15}
ggplot(filtered_qbdata, aes(x = Efficiency, y = Grade)) +
  geom_point() +
  geom_text(aes(label = Player), nudge_x = 0, nudge_y = -0.05) +
  facet_wrap(. ~ Team) +
  labs(x = "Efficiency", y = "Grade", title = "Quarterback Performance by Team")
```


## Radar Charts
```{r, fig.width=15, fig.height=12}
max <- data.frame(
  CmpPct = 0.705,
  YdsperGame = 300, #325
  TDperGame = 3,
  TOperGame = 0,
  BadThrperGame = 0,
  OnTgtperGame = 30,
  PktTime = 3.0 #3.2
)

min <- data.frame(
  CmpPct = 0.6,
  YdsperGame = 0,
  TDperGame = 0,
  TOperGame = 2.5,
  BadThrperGame = 7,
  OnTgtperGame = 0,
  PktTime = 0
)

radar_data <- filtered_qbdata %>%
  filter(ID %in% c(1, 2)) %>%
  select(CmpPct, YdsperGame, TDperGame, TOperGame, BadThrperGame, OnTgtperGame, PktTime)

radar_data <- bind_rows(max, min, radar_data)

radarchart(radar_data,
           vlabels = c("Completion %", "Yards/Game", "TD/Game", "Turnovers/Game", "Bad Throws/Game", "    Throws on \n Target/Game", "Pocket Time"),
           vlcex = 1.5, # Size of labels
           cglwd = 2, # Grid line width
           cglcol="black", # Grid line color
           pcol = "black", # Polygon Outline color
           pfcol = c(alpha("#FB4F14", 0.75), alpha("black", 0.5)), # Inside color, transparency
           plwd = 3, # Polygon Line width
           axistype = 2,
           palcex = 1.5,
           )

mtext(side = 3, line = 1, at = 0, cex = 3, "Joe Burrow vs Jake Browning Evaluation Metrics", font = 2)

legend(-1.75,-1,
       legend=c("Joe Burrow","Jake Browning"),
       pch=c(19,19),
       col=c("black","#FB4F14"))
```
