---
title: "Olympic History Data Analysis"
author: "Andre Contreras"
date: "2024-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Olympics Project

Does the economical stability of a country affect the number of athletes it sends to the olympics and the number of medals it wins?

Does hosting the olympics correlate to winning more medals that year?

### Part I

Below are the results I found for the first question, along with the code I wrote to filter and manipulate the data so I can visualize it in a more effective manner.

```{r, message=FALSE}
athlete_events <- read_csv(
   file = 'athlete_events.csv',
   col_types = cols(ID = 'i', Age = 'i', Height = 'i', Year = 'i')
)

nearest_year <- function(olympics_year) {
  gapminder_year <- seq(1952, 2007, by = 5)
  nearest_year <- gapminder_year[which.min(abs(olympics_year - gapminder_year))]
  return(nearest_year)
}

olympics_data <-
  athlete_events %>%
  filter(!is.na(Medal)) %>%
  count(Games, Event, NOC, Medal, Team, Year, Name) %>%
  mutate(year = nearest_year(Year))

country_money <- gapminder %>%
  group_by(country) %>%
  select(country, year, gdpPercap)


athletes_by_country_year <- olympics_data %>%
  group_by(Team, Year) %>%
  summarise(Total_Athletes = n_distinct(Name), .groups = 'drop')

medals_by_country_year <- olympics_data %>%
  group_by(Team, Year) %>%
  summarise(Total_Medals = n_distinct(Medal), .groups = 'drop')

joined_data_athletes <- inner_join(athletes_by_country_year, country_money, by = c("Year" = "year", "Team" = "country")) %>%
  filter(!is.na(gdpPercap))

joined_data_medals <- inner_join(medals_by_country_year, country_money, by = c("Year" = "year", "Team" = "country")) %>%
  filter(!is.na(gdpPercap))

ggplot(joined_data_athletes, aes(x = gdpPercap , y = Total_Athletes)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = FALSE) + #Plotting the athlete correlation
  labs(title = "Number of Athletes vs Country's GDP Per Capita", x = "GDP Per Capita ($)", y = "Number of Athletes Country Sends")

ggplot(joined_data_medals, aes(x = gdpPercap , y = Total_Medals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + #lm creates a smooth line to show a clear representation
    labs(title = "Medals Won vs Country's GDP Per Capita", x = "GDP Per Capita ($)", y = "Medals Won")
```

> As we can see, there is in fact a positive correlation between a country’s gdp per capita and the number of medals and athletes a country has. This means that the higher the gdp is, the more medals it wins and more athletes it sends to the Olympics.


### Part II

```{r, include=FALSE}
# Importing relevant data sets
athlete_events  = read.csv("athlete_events.csv")
host_cities = read.csv("host_cities.csv")

# Renaming the 'city' variable in host_cities to 'City' so I can join the 2 datasets with key "City"
host_cities <- rename(host_cities, "City" = "city")

data <- full_join(athlete_events, host_cities, by = "City")
```

For the second question... I started by joining data sets together and creating a function that will filter the joint dataset for each country and in each of the seasons: determine whether they hosted or not. The function also displays a plot to compare the amount of medals that country won when they hosted vs when they did not. We will then compare and draw reasonable conclusions by creating a histogram containing the average number of medals all countries combined have won when they host vs in the competitions before.

```{r}
generate_country_medals_plot <- function(country_code, country_name, summer_hosts, winter_hosts) {
 
  # SUMMER
  summer_plot <- NULL
 
  if (length(summer_hosts) > 0) {
    summer_medals <- data %>%
      filter(NOC == country_code & !is.na(Medal) & Season == "Summer" & Year %in% c(1896:2016)) %>%
      distinct(Year, Event) %>%
      group_by(Year) %>%
      summarise(Medal_Count = n())
   
    summer_medals$Host <- ifelse(summer_medals$Year %in% summer_hosts, "Hosted", "Not Hosted")
   
    summer_plot <- ggplot(summer_medals, aes(x = Year, y = Medal_Count, fill = Host)) +
      geom_bar(stat = "identity", position = "dodge") +
      geom_text(aes(label = Year), vjust = -0.5, position = position_dodge(width = 0.9)) +
      labs(x = "Year", y = "Medals", title = paste("Summer Olympic Medals won by", country_name))
  }
 
  # WINTER
  winter_plot <- NULL
 
  if (length(winter_hosts) > 0) {
    winter_medals <- data %>%
      filter(NOC == country_code & !is.na(Medal) & Season == "Winter" & Year %in% c(1896:2016)) %>%
      distinct(Year, Event) %>%
      group_by(Year) %>%
      summarise(Medal_Count = n())
   
    winter_medals$Host <- ifelse(winter_medals$Year %in% winter_hosts, "Hosted", "Not Hosted")
   
    winter_plot <- ggplot(winter_medals, aes(x = Year, y = Medal_Count, fill = Host)) +
      geom_bar(stat = "identity", position = "dodge") +
      geom_text(aes(label = Year), position = position_dodge(width = 0.9)) +
      labs(x = "Year", y = "Medals", title = paste("Winter Olympic Medals won by", country_name))
  }
 
  list(summer_plot = summer_plot, winter_plot = winter_plot)
}
```


```{r, comment="", echo=FALSE, results='hide', fig.height=3, fig.width=5}
usa_plots <- generate_country_medals_plot("USA", "United States", c(1904, 1932, 1984, 1996), c(1932, 1960, 1980, 2002))
usa_plots

australia_plots <- generate_country_medals_plot("AUS", "Australia", c(1956, 2000), numeric(0))
australia_plots

austria_plots <- generate_country_medals_plot("AUT", "Austria", numeric(0), c(1964, 1976))
austria_plots

belgium_plots <- generate_country_medals_plot("BEL", "Belgium", 1920, numeric(0))
belgium_plots

brazil_plots <- generate_country_medals_plot("BRA", "Brazil", 2016, numeric(0))
brazil_plots

canada_plots <- generate_country_medals_plot("CAN", "Canada", 1976, c(1988, 2010))
canada_plots

china_plots <- generate_country_medals_plot("CHN", "China", 2008, numeric(0))
china_plots

finland_plots <- generate_country_medals_plot("FIN", "Finland", 1952, numeric(0))
finland_plots

france_plots <- generate_country_medals_plot("FRA", "France", c(1900, 1924), c(1924, 1968, 1992))
france_plots

germany_plots <- generate_country_medals_plot(c("GER", "GDR"), "Germany", c(1936, 1972), c(1936, 1972))
germany_plots

greece_plots <- generate_country_medals_plot("GRE", "Greece", c(1896, 2004), numeric(0))
greece_plots

italy_plots <- generate_country_medals_plot("ITA", "Italy", 1960, c(1956, 2006))
italy_plots

japan_plots <- generate_country_medals_plot("JPN", "Japan", 1964, c(1972, 1998))
japan_plots

mexico_plots <- generate_country_medals_plot("MEX", "Mexico", 1968, numeric(0))
mexico_plots

netherlands_plots <- generate_country_medals_plot("NED", "Netherlands", 1928, numeric(0))
netherlands_plots

norway_plots <- generate_country_medals_plot("NOR", "Norway", numeric(0), c(1952, 1994))
norway_plots

russia_plots <- generate_country_medals_plot(c("URS", "RUS"), "Russia/USSR", 1980, 2014)
russia_plots

skorea_plots <- generate_country_medals_plot("KOR", "South Korea", 1988, numeric(0))
skorea_plots

spain_plots <- generate_country_medals_plot("ESP", "Spain", 1992, numeric(0))
spain_plots

sweden_plots <- generate_country_medals_plot("SWE", "Sweden", c(1912, 1956), numeric(0))
sweden_plots

switzerland_plots <- generate_country_medals_plot("SUI", "Switzerland", numeric(0), c(1928, 1948))
switzerland_plots

uk_plots <- generate_country_medals_plot("GBR", "United Kingdom", c(1908, 1948, 2012), numeric(0))
uk_plots

yugoslavia_plots <- generate_country_medals_plot("YUG", "Yugoslavia", numeric(0), 1984)
yugoslavia_plots
```

As stated earlier, I created a histogram of the difference of medals **(by subtracting the medals won when they host minus the medals won in the olympic season directly prior)** to draw a reasonable conclusion.

```{r}
hist_data <- tibble(
  NOC = c("USA", "AUS", "AUT", "BEL", "BRA", "CAN", "CHN", "FIN", "FRA", "GER", "GRE", "ITA", "JPN", "MEX", "NED", "NOR", "RUS", "KOR", "ESP", "SWE", "SUI", "GBR", "YUG"),
  Medals_Won_Host = c(451, 83, 13, 33, 18, 39, 85, 22, 109, 102, 45, 47, 36, 9, 22, 29, 130, 30, 22, 64, 9, 165, 1),
  Medals_Won_Year_Before = c(265, 49, 8, 5, 17, 29, 55, 19, 52, 39, 16, 33, 22, 1, 11, 18, 84, 18, 4, 53, 5, 76, 3))

new <- hist_data %>%
  mutate(Distribution_difference = Medals_Won_Host - Medals_Won_Year_Before)

hist(new$Distribution_difference, xlab = "Difference between Medals (Host Season - Season Before)", main = "Histogram of the Distribution Difference", col = "lightgreen")
```

> We can see there is a positive host effect country on the amount of medals won when a country hosts the olympics vs when they don't because there is an overall positive difference.