---
title: "My Portfolio"
author: "Andre Contreras"
date: "2023-2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 2011 Masters Golf Tournament Project

 The plot below is a line graph I created visualizing the summary of the Masters 2011 Pro Golf Tournament, along with the performance of each golfer and the overall winner of the comptetition (Charl Schwartzel). 

```{r, include=FALSE}
# For Golf:
library(tidyverse)
load(url(
  "https://www.stat.osu.edu/~vqv/4194/data/masters2011-untidy.rda"
))

# For NFL:
library(readxl)
library(ggplot2)
library(ggrepel)

# For Olympics:
library(gapminder)
library(countrycode)
library(dplyr)
```

```{r, fig.width=20, fig.height=12, warning=FALSE, message=FALSE}
# Binds the rows of round1, round2,round3, and round4; specifies the name of the new column that will be created to store the source of each row in the resulting data frame, rounds
rounds <- bind_rows(round1, round2, round3, round4, .id = "round")


scorecard <- rounds %>% 
  pivot_longer(cols = "1":"18", names_to = "hole", values_to = "score") %>% 
  mutate(round = as.integer(round), hole = as.integer(hole), score = as.integer(score))


performance <- scorecard %>%
  left_join(course, by = "hole") %>%
  mutate(difference_to_par = score - par) %>%
  group_by(player) %>%
  mutate(cumulative_to_par = cumsum(difference_to_par)) %>%
  ungroup() %>%
  select(player, round, hole, difference_to_par, cumulative_to_par)

winner <- performance %>%
  filter(round == 4) %>%
  top_n(1, wt = -cumulative_to_par)
  
ggplot(performance) +
  geom_line(aes(x = hole, y = cumulative_to_par, col = player)) +
  facet_grid(. ~ round, labeller = labeller(round = c("1" = "round 1", "2" = "round 2", "3" = "round 3", "4" = "round 4"))) +
  geom_label(aes(x = hole - 4, y = cumulative_to_par, label = player), data = winner) +
  labs(title = "Performance Summary: Masters 2011 Pro Golf Tournament", x = "Hole", y = "Cumulative Score to Par", caption = "Overall winner: Charl Schwartzel") +
  theme_minimal() +
  theme(legend.position = "right", legend.key.size = unit(c(1, 1), "cm"), legend.text = element_text(size = 10))

```


# ______________________________________________________________________________

# NFL 2023 QB Performance Project


### Overall QB Performance:

 I created a spreadsheet of all the NFL Quarterbacks that played a minimum of one game in the 2023-2024 season and compiled all of their advanced statistics. I then developed a formula in Excel aiming to grade their efficiency, MVP rating, and overall grade, and transferred the file into R to visualize my findings. The name displayed is my MVP according to my grading system and efficiency formula

```{r, fig.width=12, warning=FALSE, message=FALSE}
qbdata <- read_excel("C:\\Users\\ajcon\\Downloads\\Portfolio\\qbdata.xlsx")

mvp <- qbdata %>% 
  top_n(1, qbdata$Grade*qbdata$Efficiency)

ggplot(qbdata, aes(x = Efficiency, y = Grade, col = QB)) +
  geom_point() +
  geom_label_repel(data = mvp, aes(label = QB), nudge_x = -0.001, nudge_y = -0.05, segment.color = "black", segment.size = 0.5) +
  labs(x = "Weighted Efficiency Scale", y = "Grading Scale", title = "Overall QB Performance 2023 NFL Season") +
  theme_minimal() +
  theme(legend.position = "right")

```


### QB Performance by Team:

```{r, fig.width=20, fig.height=15}
ggplot(qbdata, aes(x = Efficiency, y = Grade)) +
  geom_point() +
  geom_text(aes(label = QB), nudge_x = 0, nudge_y = -0.05) +
  facet_wrap(. ~ Team) +
  labs(x = "Efficiency", y = "Grade", title = "Quarterback Performance by Team")
```

# ______________________________________________________________________________

# A Data-led Look into the History of the Olympics

 After gaining access to mulitple datasets of the Olympics containing every instance throughout every competition since the inaugural season back in 1896 (Greece) up until the 2016 Games in Brazil, I decided my free time would be well spent answering a couple of questions I, like many others (I think), have been wondering:

> 1. Does the economical stability of a country affect the number of athletes it sends to the olympics and the number of medals it wins?

> 2. Does hosting the olympics correlate to winning more medals that year?

### Part I

Below are the results I found for the first question, along with the code I wrote to filter and manipulate the data so I can visualize it in a more effective manner.

```{r, message=FALSE}
athlete_events <- read_csv(
   file = 'athlete_events.csv',
   col_types = cols(ID = 'i', Age = 'i', Height = 'i', Year = 'i')
)

nearest_year <- function(olympics_year) {
  gapminder_year <- seq(1952, 2007, by = 5)
  nearest_year <- gapminder_year[which.min(abs(olympics_year - gapminder_year))]
  return(nearest_year)
}

olympics_data <-
  athlete_events %>%
  filter(!is.na(Medal)) %>%
  count(Games, Event, NOC, Medal, Team, Year, Name) %>%
  mutate(year = nearest_year(Year))

country_money <- gapminder %>%
  group_by(country) %>%
  select(country, year, gdpPercap)


athletes_by_country_year <- olympics_data %>%
  group_by(Team, Year) %>%
  summarise(Total_Athletes = n_distinct(Name), .groups = 'drop')

medals_by_country_year <- olympics_data %>%
  group_by(Team, Year) %>%
  summarise(Total_Medals = n_distinct(Medal), .groups = 'drop')

joined_data_athletes <- inner_join(athletes_by_country_year, country_money, by = c("Year" = "year", "Team" = "country")) %>%
  filter(!is.na(gdpPercap))

joined_data_medals <- inner_join(medals_by_country_year, country_money, by = c("Year" = "year", "Team" = "country")) %>%
  filter(!is.na(gdpPercap))

ggplot(joined_data_athletes, aes(x = gdpPercap , y = Total_Athletes)) +
  geom_point() +
  geom_smooth(method = "lm", color = "red", se = FALSE) + #Plotting the athlete correlation
  labs(title = "Number of Athletes vs Country's GDP Per Capita", x = "GDP Per Capita ($)", y = "Number of Athletes Country Sends")

ggplot(joined_data_medals, aes(x = gdpPercap , y = Total_Medals)) +
  geom_point() +
  geom_smooth(method = "lm", color = "blue", se = FALSE) + #lm creates a smooth line to show a clear representation
    labs(title = "Medals Won vs Country's GDP Per Capita", x = "GDP Per Capita ($)", y = "Medals Won")
```

> As we can see, there is in fact a positive correlation between a country’s gdp per capita and the number of medals and athletes a country has. This means that the higher the gdp is, the more medals it wins and more athletes it sends to the Olympics.


### Part II

```{r, include=FALSE}
# Importing relevant data sets
athlete_events  = read.csv("athlete_events.csv")
host_cities = read.csv("host_cities.csv")

# Renaming the 'city' variable in host_cities to 'City' so I can join the 2 datasets with key "City"
host_cities <- rename(host_cities, "City" = "city")

data <- full_join(athlete_events, host_cities, by = "City")
```

For the second question... I started by joining data sets together and creating a function that will filter the joint dataset for each country and in each of the seasons: determine whether they hosted or not. The function also displays a plot to compare the amount of medals that country won when they hosted vs when they did not. We will then compare and draw reasonable conclusions by creating a histogram containing the average number of medals all countries combined have won when they host vs in the competitions before.

```{r}
generate_country_medals_plot <- function(country_code, country_name, summer_hosts, winter_hosts) {
 
  # SUMMER
  summer_plot <- NULL
 
  if (length(summer_hosts) > 0) {
    summer_medals <- data %>%
      filter(NOC == country_code & !is.na(Medal) & Season == "Summer" & Year %in% c(1896:2016)) %>%
      distinct(Year, Event) %>%
      group_by(Year) %>%
      summarise(Medal_Count = n())
   
    summer_medals$Host <- ifelse(summer_medals$Year %in% summer_hosts, "Hosted", "Not Hosted")
   
    summer_plot <- ggplot(summer_medals, aes(x = Year, y = Medal_Count, fill = Host)) +
      geom_bar(stat = "identity", position = "dodge") +
      geom_text(aes(label = Year), vjust = -0.5, position = position_dodge(width = 0.9)) +
      labs(x = "Year", y = "Medals", title = paste("Summer Olympic Medals won by", country_name))
  }
 
  # WINTER
  winter_plot <- NULL
 
  if (length(winter_hosts) > 0) {
    winter_medals <- data %>%
      filter(NOC == country_code & !is.na(Medal) & Season == "Winter" & Year %in% c(1896:2016)) %>%
      distinct(Year, Event) %>%
      group_by(Year) %>%
      summarise(Medal_Count = n())
   
    winter_medals$Host <- ifelse(winter_medals$Year %in% winter_hosts, "Hosted", "Not Hosted")
   
    winter_plot <- ggplot(winter_medals, aes(x = Year, y = Medal_Count, fill = Host)) +
      geom_bar(stat = "identity", position = "dodge") +
      geom_text(aes(label = Year), position = position_dodge(width = 0.9)) +
      labs(x = "Year", y = "Medals", title = paste("Winter Olympic Medals won by", country_name))
  }
 
  list(summer_plot = summer_plot, winter_plot = winter_plot)
}
```


```{r, comment="", echo=FALSE, results='hide'}
usa_plots <- generate_country_medals_plot("USA", "United States", c(1904, 1932, 1984, 1996), c(1932, 1960, 1980, 2002))
usa_plots

australia_plots <- generate_country_medals_plot("AUS", "Australia", c(1956, 2000), numeric(0))
australia_plots

austria_plots <- generate_country_medals_plot("AUT", "Austria", numeric(0), c(1964, 1976))
austria_plots

belgium_plots <- generate_country_medals_plot("BEL", "Belgium", 1920, numeric(0))
belgium_plots

brazil_plots <- generate_country_medals_plot("BRA", "Brazil", 2016, numeric(0))
brazil_plots

canada_plots <- generate_country_medals_plot("CAN", "Canada", 1976, c(1988, 2010))
canada_plots

china_plots <- generate_country_medals_plot("CHN", "China", 2008, numeric(0))
china_plots

finland_plots <- generate_country_medals_plot("FIN", "Finland", 1952, numeric(0))
finland_plots

france_plots <- generate_country_medals_plot("FRA", "France", c(1900, 1924), c(1924, 1968, 1992))
france_plots

germany_plots <- generate_country_medals_plot(c("GER", "GDR"), "Germany", c(1936, 1972), c(1936, 1972))
germany_plots

greece_plots <- generate_country_medals_plot("GRE", "Greece", c(1896, 2004), numeric(0))
greece_plots

italy_plots <- generate_country_medals_plot("ITA", "Italy", 1960, c(1956, 2006))
italy_plots

japan_plots <- generate_country_medals_plot("JPN", "Japan", 1964, c(1972, 1998))
japan_plots

mexico_plots <- generate_country_medals_plot("MEX", "Mexico", 1968, numeric(0))
mexico_plots

netherlands_plots <- generate_country_medals_plot("NED", "Netherlands", 1928, numeric(0))
netherlands_plots

norway_plots <- generate_country_medals_plot("NOR", "Norway", numeric(0), c(1952, 1994))
norway_plots

russia_plots <- generate_country_medals_plot(c("URS", "RUS"), "Russia/USSR", 1980, 2014)
russia_plots

skorea_plots <- generate_country_medals_plot("KOR", "South Korea", 1988, numeric(0))
skorea_plots

spain_plots <- generate_country_medals_plot("ESP", "Spain", 1992, numeric(0))
spain_plots

sweden_plots <- generate_country_medals_plot("SWE", "Sweden", c(1912, 1956), numeric(0))
sweden_plots

switzerland_plots <- generate_country_medals_plot("SUI", "Switzerland", numeric(0), c(1928, 1948))
switzerland_plots

uk_plots <- generate_country_medals_plot("GBR", "United Kingdom", c(1908, 1948, 2012), numeric(0))
uk_plots

yugoslavia_plots <- generate_country_medals_plot("YUG", "Yugoslavia", numeric(0), 1984)
yugoslavia_plots
```

As stated earlier, I created a histogram of the difference of medals **(by subtracting the medals won when they host minus the medals won in the olympic season directly prior)** to draw a reasonable conclusion.

```{r}
hist_data <- tibble(
  NOC = c("USA", "AUS", "AUT", "BEL", "BRA", "CAN", "CHN", "FIN", "FRA", "GER", "GRE", "ITA", "JPN", "MEX", "NED", "NOR", "RUS", "KOR", "ESP", "SWE", "SUI", "GBR", "YUG"),
  Medals_Won_Host = c(451, 83, 13, 33, 18, 39, 85, 22, 109, 102, 45, 47, 36, 9, 22, 29, 130, 30, 22, 64, 9, 165, 1),
  Medals_Won_Year_Before = c(265, 49, 8, 5, 17, 29, 55, 19, 52, 39, 16, 33, 22, 1, 11, 18, 84, 18, 4, 53, 5, 76, 3))

new <- hist_data %>%
  mutate(Distribution_difference = Medals_Won_Host - Medals_Won_Year_Before)

hist(new$Distribution_difference, xlab = "Difference between Medals (Host Season - Season Before)", main = "Histogram of the Distribution Difference", col = "lightgreen")
```

> We can see there is a positive host effect country on the amount of medals won when a country hosts the olympics vs when they don't because there is an overall positive difference.

# ______________________________________________________________________________

#### I am also currently in the process of completing my fourth project - Analyzing the passing networks of the undefeated 2003-2004 Arsenal FC soccer team to use Graph Theory concepts and techniques in hopes of optimizing strategies that will lead to better results and more wins in present day soccer.

#### I am also planning on undertaking a Machine Learning Project later this summer!