---
title: "Premier League Project"
author: "Andre Contreras"
date: "2024-06-13"
output: 
  prettydoc::html_pretty:
    theme: tactile
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(ggimage)
library(tidyr)
library(caret)

images <- data.frame(
  Team = c("Manchester City", "Liverpool", "Leicester City", "Chelsea", "Manchester Utd", "Arsenal", "Blackburn"),
  Logo = c(
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Manchester-City-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Liverpool-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Leicester-City-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Chelsea-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Manchester-United-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Arsenal-Logo.png",
    "C:/Users/ajcon/Downloads/Portfolio/Premier League Project/team_logos/Blackburn-Rovers-Logo.png"
  )
)

pldata <- read_csv("C:/Users/ajcon/Downloads/Portfolio/Premier League Project/premier-league-table.csv")

winners <- pldata %>% 
    filter(Rk == 1)

winners <- left_join(winners, images, by = "Team")

team_colors <- c(
  "Manchester City" = "#6CABDD",
  "Arsenal" = "#EF0107",
  "Liverpool" = "#C8102E",
  "Aston Villa" = "#95BFE5",
  "Tottenham" = "#132257",
  "Chelsea" = "#034694",
  "Newcastle Utd" = "#241F20",
  "Manchester Utd" = "#DA291C",
  "West Ham" = "#7A263A",
  "Crystal Palace" = "#1B458F",
  "Brighton" = "#0057B8",
  "Bournemouth" = "#DA291C",
  "Fulham" = "#000000",
  "Wolves" = "#FDB913",
  "Everton" = "#003399",
  "Brentford" = "#E30613",
  "Nottingham Forest" = "#DD0000",
  "Luton Town" = "#F78F1E",
  "Burnley" = "#6C1D45",
  "Sheffield Utd" = "#EE2737",
  "Leeds United" = "#FFCD00",
  "Leicester City" = "#003090",
  "Southampton" = "#D71920",
  "Norwich City" = "#00A650",
  "Watford" = "#11210C",
  "West Brom" = "#122F67",
  "Cardiff City" = "#1B458F",
  "Huddersfield" = "#0073CF",
  "Stoke City" = "#E03A3E",
  "Swansea City" = "#121212",
  "Hull City" = "#F18A01",
  "Middlesbrough" = "#E03A3E",
  "Sunderland" = "#EB172B",
  "QPR" = "#2C2A29",
  "Reading" = "#0B3669",
  "Wigan Athletic" = "#0066CC",
  "Blackburn" = "#1C2C3B",
  "Bolton" = "#112B54",
  "Birmingham City" = "#003366",
  "Blackpool" = "#EE3524",
  "Portsmouth" = "#0034A3",
  "Derby County" = "#0066CC",
  "Charlton Ath" = "#EF2B2D",
  "Ipswich Town" = "#00ADEF",
  "Bradford City" = "#6d767a",
  "Coventry City" = "#6BA539",
  "Sheffield Weds" = "#0033A0",
  "Wimbledon" = "#F58220",
  "Barnsley" = "#EE2737",
  "Oldham Athletic" = "#0033A0",
  "Swindon Town" = "#EA5B0C"
)

```


## 2024 Premier League Standings by Matchweek
```{r echo=FALSE, fig.height=15, fig.width=35, message=FALSE, warning=FALSE}
pl2024 <- read_csv("C:/Users/ajcon/Downloads/Portfolio/Premier League Project/premier-league-2024-standings.csv")

label_data <- pl2024 %>%
  group_by(Team) %>%
  filter(Matchweek == max(Matchweek)) %>%
  ungroup()

ggplot(pl2024) +
  geom_line(aes(x = Matchweek, y = Standing, col = Team), size = 1.5, show.legend = FALSE) +
  labs(x = "Matchweek", y = "Standing", title = "2023-2024 Premier League Standings by Matchweek") +
  geom_text(data = label_data, aes(x = Matchweek, y = Standing, label = Team, col = Team), hjust = -0.3, vjust = 0.5, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(1, 38, by = 1)) +
  scale_y_reverse(breaks = seq(1, 20, by = 1)) +
  scale_color_manual(values = team_colors) +
  theme(
    plot.title = element_text(size = 60, face = "bold"),
    axis.title.x = element_text(size = 40),
    axis.title.y = element_text(size = 40),
    axis.text.x = element_text(size = 25),
    axis.text.y = element_text(size = 25))
```


# History of Premier League Table
```{r, warning=FALSE, message=FALSE, warning=FALSE, fig.height = 20, fig.width = 35, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE) # Caching it so it doesn't have to rerun when knitting. it has everything stored making it knit faster

pldata <- pldata %>% 
  filter(Rk < 21)

label_data <- pldata %>% 
    group_by(Team) %>% 
    filter(Season_End_Year == max(Season_End_Year)) %>% 
    ungroup()
  

# Plot of all teams
ggplot(pldata) +
    geom_line(aes(x = Season_End_Year, y = Rk, col = Team), size = 1.5, show.legend = FALSE, linetype = "solid") +
    geom_text(data = label_data, aes(x = Season_End_Year, y = Rk, label = Team, col = Team), hjust = -0.3, vjust = 0.5, show.legend = FALSE) +
    geom_image(data = winners, aes(x = Season_End_Year, y = 0.5, image = Logo), size = 0.05) + # Adjust size as needed
    labs(x = "Year", y = "Standing", title = "History of the Premier League Table") +
    scale_x_continuous(breaks = seq(1993, 2024, by = 1)) +
    scale_y_reverse(breaks = seq(1, 20, by = 1)) +
    scale_color_manual(values = team_colors) +  
    theme(
    plot.title = element_text(size = 75, face = "bold"),
    axis.title.x = element_text(size = 30),
    axis.title.y = element_text(size = 30),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15)
  )
```

#### Average Finishing Position of each club since the establishment of the Premier League (1993)
```{r, echo=FALSE}
plfinish <- pldata %>% 
  group_by(Team) %>% 
  summarise(Average_finishing_position = round(mean(Rk), 3)) %>% 
  arrange(-desc(Average_finishing_position))

plfinish
```

#### Total number of wins for each club since establishment of the Premier League
```{r, echo=FALSE}
plwins <- pldata %>% 
  group_by(Team) %>% 
  summarise(Total_Wins = sum(W)) %>% 
  arrange(desc(Total_Wins))

plwins
```

#### Total number of points for each club since the establishment of the Premier League
```{r, echo=FALSE}
plpoints <- pldata %>% 
  group_by(Team) %>% 
  summarise(Total_Points = sum(Pts)) %>% 
  arrange(desc(Total_Points))

plpoints
```


## Premier League Champions by the Years
```{r, fig.width=35, fig.height=20, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE) # Caching it so it doesn't have to rerun when knitting. it has everything stored making it knit faster

plchampions <- pldata %>% 
  filter(Team %in% c("Arsenal", "Manchester City", "Manchester Utd", "Chelsea", "Liverpool", "Blackburn", "Leicester City"))

plchampions <- plchampions %>% 
  complete(Team, Season_End_Year) # USEFUL - filled in all the teams years where they weren't in the premier league bc of relegation with NA (tidyr package)

label_data <- plchampions %>% 
  group_by(Team) %>% 
    filter(Season_End_Year == max(Season_End_Year)) %>% 
    ungroup()

label_data <- bind_rows(pldata %>% filter(Team == "Blackburn" & Season_End_Year == 2012), pldata %>% filter(Team == "Leicester City" & Season_End_Year == 2023), label_data)
  
# Plot of only Premier League Champions
ggplot(plchampions) +
    geom_line(aes(x = Season_End_Year, y = Rk, col = Team), size = 1.5, show.legend = FALSE) +
    geom_point(aes(x = Season_End_Year, y = Rk, col = Team), size = 5, show.legend = FALSE) +
    geom_text(data = label_data, aes(x = Season_End_Year, y = Rk, label = Team, col = Team), hjust = -0.3, vjust = 0.5, show.legend = FALSE) +
    geom_image(data = winners, aes(x = Season_End_Year, y = 0.5, image = Logo), size = 0.05) + # Adjust size as needed
    labs(x = "Year", y = "Standing", title = "Premier League Champions by Year") +
    scale_x_continuous(breaks = seq(1993, 2024, by = 1)) +
    scale_y_reverse(breaks = seq(1, 22, by = 1)) +
    scale_color_manual(values = team_colors) +
    theme(
    plot.title = element_text(size = 75, face = "bold"),
    axis.title.x = element_text(size = 30),
    axis.title.y = element_text(size = 30),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15))
```

```{r, fig.width=35, fig.height=5, echo=FALSE}
plot.new()

# Add a legend
legend("top",    # Position of the legend (e.g., "topright", "bottomleft", etc.)
       legend = "Indicates a team was promoted to the Premier League at the start of the season, but later relegated at the end of the season",
       pch = 19,       # Symbol type (19 is a solid circle)
       col = "black",  # Color of the symbol
       cex = 3)        # Size of the symbol
```



## Premier League Big 6 by the Years
```{r echo=FALSE, fig.height=20, fig.width=35, message=FALSE, warning=FALSE}

plbigsix <- pldata %>% 
  filter(Team %in% c("Arsenal", "Tottenham", "Manchester City", "Manchester Utd", "Chelsea", "Liverpool"))

plbigsix <- plbigsix %>% 
  complete(Team, Season_End_Year)

label_data <- plbigsix %>% 
  group_by(Team) %>% 
    filter(Season_End_Year == max(Season_End_Year)) %>% 
    ungroup()
  
# Plot of only Premier League Champions
ggplot(plbigsix) +
    geom_line(aes(x = Season_End_Year, y = Rk, col = Team), linewidth = 1.5, show.legend = FALSE) +
    geom_point(aes(x = Season_End_Year, y = Rk, col = Team), size = 5, show.legend = FALSE) +
    geom_text(data = label_data, aes(x = Season_End_Year, y = Rk, label = Team, col = Team), hjust = -0.3, vjust = 0.5, show.legend = FALSE) +
    labs(x = "Year", y = "Standing", title = "Premier League Big 6 by Year") +
    scale_x_continuous(breaks = seq(1993, 2024, by = 1)) +
    scale_y_reverse(breaks = seq(1, 20, by = 1)) +
    scale_color_manual(values = team_colors) +
    theme(
    plot.title = element_text(size = 75, face = "bold"),
    axis.title.x = element_text(size = 50),
    axis.title.y = element_text(size = 50),
    axis.text.x = element_text(size = 25),
    axis.text.y = element_text(size = 25))
```

```{r, fig.width=35, fig.height=5, echo=FALSE}
plot.new()

# Add a legend
legend("top",    # Position of the legend (e.g., "topright", "bottomleft", etc.)
       legend = "Indicates a team was promoted to the Premier League at the start of the season, but later relegated at the end of the season",
       pch = 19,       # Symbol type (19 is a solid circle)
       col = "black",  # Color of the symbol
       cex = 3)        # Size of the symbol
```


# Model to predict Premier League Champion
```{r, warning=FALSE, message=FALSE}
set.seed(100)
# Step 1: Feature Engineering
pldata <- pldata %>%
  select(c(1:11)) %>%
  mutate(isChampion = ifelse(Rk == 1, 1, 0)) %>% 
  group_by(Team) %>% 
  mutate(Experience = sum(MP)/38) %>% 
  ungroup() %>% 
  filter(Experience >= 3)

# Step 2: Split into 2 sets
index <- createDataPartition(pldata$isChampion, p = 0.6, list = FALSE)
training <- pldata[index, ]
testing <- pldata[-index, ]

# Step 3: Train Model
model <- glm(isChampion ~ W + D + L + GF + GA + GD + Pts, data = training, family = binomial)


# Step 4: Evaluate Model
probability <- round(predict(model, newdata = testing, type = "response"), 3)

prediction <- ifelse(probability > 0.5, 1, 0)

new_pldata <- bind_cols(testing, probability, prediction)

results <- data.frame("Probability To Win PL" = probability, "Prediction" = prediction, "Actual" = testing$isChampion)
results

table("Prediction" = prediction, "Actual" = testing$isChampion)

# Wins plot
ggplot(testing, aes(x = W, y = isChampion)) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) +
  labs(y = "Probability of Winning Premier League", x = "Number of Wins", title = "Probability of becoming Champion of Premier League given Number of Wins") +
  scale_x_continuous(breaks = seq(1, 40, by = 5), limits = c(1, 38)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1))

# Points plot
ggplot(testing, aes(x = Pts, y = isChampion)) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) +
  labs(y = "Probability of Winning Premier League", x = "Points", title = "Probability of becoming Champion of Premier League given Points") +
  scale_x_continuous(breaks = seq(1, 110, by = 10)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1))

# Goals for plot
ggplot(testing, aes(x = GF, y = isChampion)) +
  geom_point() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = binomial)) +
  labs(y = "Probability of Winning Premier League", x = "Goals For", title = "Probability of becoming Champion of Premier League given Goals scored") +
  scale_x_continuous(breaks = seq(1, 100, by = 10)) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.1))
```

